---
title: "Desmognathus study data, methods, and figures"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
    df_print: paged
    theme: readable
    highlight: tango
---

<style>
body { max-width: 1500px; }
.callout { border-left: 5px solid #0072B2; background: #f3f8fb; padding: 0.8em 1em; margin: 1em 0; }
.warning { border-left-color: #D55E00; background: #fff7f2; }
.boundary { border-left-color: #E69F00; background: #fffaf0; }
code { white-space: pre-wrap; }
.figure img { border: 1px solid #D9DEE2; }
</style>

```{r report-setup}
suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(ggrepel)
  library(knitr)
  library(patchwork)
  library(readr)
  library(rmarkdown)
  library(scales)
  library(tidyr)
})

ROOT <- normalizePath(getwd(), mustWork = TRUE)
DATA <- file.path(ROOT, "Publication", "datasets")
FIGURES <- file.path(ROOT, "Publication", "figures")
REPORT_SOURCE <- file.path(ROOT, "Publication", "Desmognathus_study_data.Rmd")

source(file.path(ROOT, "scripts", "publication", "gbe_theme.R"))
knitr::read_chunk(file.path(ROOT, "scripts", "publication", "build_gbe_figures.R"))
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  out.width = "100%"
)
options(knitr.kable.NA = "")

manifest <- read_csv(
  file.path(DATA, "DATASET_MANIFEST.csv"),
  show_col_types = FALSE,
  progress = FALSE
)

manifest_row <- function(supplement_id) {
  row <- manifest %>% filter(.data$supplement_id == .env$supplement_id)
  if (nrow(row) != 1L) stop("Expected one manifest row for ", supplement_id)
  row
}

read_supp <- function(supplement_id) {
  row <- manifest_row(supplement_id)
  read_csv(
    file.path(DATA, row$filename[[1]]),
    show_col_types = FALSE,
    progress = FALSE
  )
}

show_table <- function(data, rows = 10L, columns = 12L) {
  rmarkdown::paged_table(
    data,
    options = list(
      rows.print = min(rows, nrow(data)),
      cols.print = min(columns, ncol(data))
    )
  )
}

sha256_file <- function(path) {
  output <- system2("sha256sum", path, stdout = TRUE, stderr = TRUE)
  if (length(output) != 1L) stop("Could not hash ", path)
  strsplit(output, "[[:space:]]+")[[1]][[1]]
}

species_label <- function(x) {
  text <- as.character(x)
  ifelse(grepl("^D\\. ", text), text, paste("D.", text))
}

figure_path <- function(filename) {
  path <- file.path(FIGURES, filename)
  if (!file.exists(path)) stop("Missing figure: ", path, ". Run `make figures`.")
  path
}

stopifnot(
  nrow(manifest) == 40L,
  identical(manifest$supplement_id, sprintf("S%02d", seq_len(40)))
)
```

<div class="callout">
This is a **read-only researcher-facing view** over the controlled study
release. The CSV files and their hashes remain the data authority. Code chunks
make the filters and equations visible; the report does not overwrite a source
table, review decision, or publication dataset.
</div>

<div class="callout boundary">
**Claim boundaries.** Relative nuclear IOD is an image-derived relative
phenotype, not an independently validated absolute genome size. The separate
*D. fuscus* scale uses an assembly-derived conditional anchor of 16.462167689
pg/1C (`16.1 Gbp / 0.978 Gbp per pg`). The LTR terminal:internal statistic is a
mapping/deletion-footprint proxy, not a direct ectopic-recombination, solo-LTR,
or DNA-loss rate. Cell and nucleus summaries describe the reviewed upper tail
of observed erythrocytes and their intervals are conditional on the observed
animals and images. Current path analyses are exploratory and do not orient a
unique causal arrow.
</div>

# Study report {.tabset .tabset-fade}

## Guide and dataset index

### What is included

- **TE34:** 34 species with accession-linked genomic resources.
- **Path24:** 24 species with finalized cell, nucleus, and relative-IOD data.
- **overlap21:** the exact 21-species intersection; no TE or phenotype values
  are imputed.
- **S01–S40:** controlled CSV supplements with source/output SHA-256 hashes.
- **Figures:** the exact current ggplot figures built by
  [`scripts/publication/build_gbe_figures.R`](../scripts/publication/build_gbe_figures.R).

The report shows compact species-level tables directly. Large object- and
element-level tables remain downloadable CSVs so the HTML stays responsive.

```{r dataset-index}
manifest %>%
  select(
    supplement_id,
    analysis_section,
    title,
    release_status,
    manuscript_use,
    rows,
    columns,
    filename,
    output_sha256
  ) %>%
  show_table(rows = 12, columns = 9)
```

### Package versions

```{r package-versions}
tibble(
  package = c(
    "R", "rmarkdown", "knitr", "readr", "dplyr", "tidyr",
    "ggplot2", "ggrepel", "patchwork", "scales"
  ),
  version = c(
    as.character(getRversion()),
    vapply(
      c(
        "rmarkdown", "knitr", "readr", "dplyr", "tidyr",
        "ggplot2", "ggrepel", "patchwork", "scales"
      ),
      function(package) as.character(packageVersion(package)),
      character(1)
    )
  )
)
```

## Study panels and genomic QC

The report checks the panel contract directly before showing the current study
design and genomic-resource figure.

```{r panel-contract}
availability <- read_supp("S01")
te_panel <- read_supp("S02")
path_panel <- read_supp("S03")
path_traits <- read_supp("S04")
assembly_quality <- read_supp("S05")
dnapipete_quality <- read_supp("S06")

overlap_species <- inner_join(
  te_panel %>% select(species),
  path_panel %>% select(species),
  by = "species"
)

stopifnot(
  nrow(te_panel) == 34L,
  n_distinct(te_panel$species) == 34L,
  nrow(path_panel) == 24L,
  n_distinct(path_panel$species) == 24L,
  nrow(overlap_species) == 21L,
  nrow(path_traits) == 24L
)

tibble(
  panel = c("TE34", "Path24", "overlap21"),
  n_species = c(nrow(te_panel), nrow(path_panel), nrow(overlap_species))
)
```

### Species availability table

```{r availability-table}
availability %>% show_table(rows = 12, columns = 12)
```

### Exact ggplot code and current figure

The following collapsed source chunk is the production ggplot code, read from
the single figure builder rather than copied into this report.

```{r figure-1, eval=FALSE}
```

```{r show-figure-1, echo=FALSE, fig.cap="Figure 1. Path24 time tree, study panels, and genomic QC."}
knitr::include_graphics(
  figure_path("Figure_1_study_design_and_genomic_qc.png"),
  rel_path = FALSE
)
```

## TE diversity and PCA

### Diversity equations and primary filter

For proportions \(p_i\) within a species:

\[
H' = -\sum_i p_i \ln(p_i)
\]

\[
D_{GS} = 1 - \sum_i p_i^2
\]

These are natural-log Shannon entropy and the Gini-Simpson index.

The paper-facing primary stratum is `te_level == "order"`. Observed richness
is retained only as an audit count, and the superfamily stratum is retained as
a sensitivity/audit view. S08 contains classified-only diversity: classified
categories are reclosed to one within each species. Unresolved aligned-base
mass remains separate in S07 as dnaPipeTE classification-accounting QC and is
not treated as a diversity category.

```{r diversity-replay}
te_diversity <- read_supp("S08")
order_composition <- read_csv(
  file.path(
    ROOT, "analyses", "01_transposable_elements", "data",
    "te_order_composition.csv"
  ),
  show_col_types = FALSE
)

primary_diversity <- te_diversity %>%
  filter(te_level == "order")

recomputed_primary <- order_composition %>%
  pivot_longer(-species, names_to = "feature", values_to = "proportion") %>%
  group_by(species) %>%
  summarise(
    observed_richness = sum(proportion > 0),
    shannon_entropy = -sum(proportion[proportion > 0] * log(proportion[proportion > 0])),
    gini_simpson = 1 - sum(proportion^2),
    .groups = "drop"
  ) %>%
  arrange(species)

observed_primary <- primary_diversity %>%
  select(species, observed_richness, shannon_entropy, gini_simpson) %>%
  arrange(species)

stopifnot(
  nrow(te_diversity) == 68L,
  nrow(primary_diversity) == 34L,
  n_distinct(te_diversity$species) == 34L,
  setequal(te_diversity$te_level, c("order", "superfamily")),
  nrow(te_diversity %>% count(te_level)) == 2L,
  all(te_diversity %>% count(te_level) %>% pull(n) == 34L),
  !any(grepl("unresolved", names(te_diversity), ignore.case = TRUE)),
  isTRUE(all.equal(recomputed_primary, observed_primary, tolerance = 1e-12))
)

primary_diversity %>%
  arrange(desc(shannon_entropy)) %>%
  show_table(rows = 12, columns = 5)
```

### PCA normalization

The PCA uses retry-averaged superfamily compositions, retains features positive
in all 34 TE species, closes each row, and applies the centered log-ratio
transformation:

\[
\operatorname{clr}(p_{ij}) = \ln(p_{ij}) - \frac{1}{D}\sum_{k=1}^{D}\ln(p_{ik})
\]

Columns are centered and an unscaled SVD is fit. There is no zero replacement
in the retained 24-superfamily matrix. CR1, Chapaev, Dada, Ginger, and Merlin
remain available for diversity but fail the ubiquitous-feature PCA gate.

```{r pca-contract}
te_scores <- read_supp("S09")
te_variance <- read_supp("S10")
te_loadings <- read_supp("S11")
feature_prevalence <- read_csv(
  file.path(
    ROOT, "analyses", "01_transposable_elements", "data",
    "te_feature_prevalence.csv"
  ),
  show_col_types = FALSE
)

retained_superfamilies <- feature_prevalence %>%
  filter(te_level == "superfamily", pca_included)
excluded_superfamilies <- feature_prevalence %>%
  filter(te_level == "superfamily", !pca_included)

stopifnot(
  nrow(te_scores) == 34L,
  nrow(te_variance) == 23L,
  nrow(te_loadings) == 24L,
  nrow(retained_superfamilies) == 24L,
  setequal(
    excluded_superfamilies$feature,
    c("CR1", "Chapaev", "Dada", "Ginger", "Merlin")
  )
)

excluded_superfamilies %>%
  select(
    feature, n_species_positive, prevalence_fraction,
    minimum_proportion, maximum_proportion, pca_exclusion_reason
  )
```

No final sensitivity analysis asks whether forcing CR1 or Chapaev into the
shared-feature PCA changes support; the report does not invent one.

### Exact ggplot code and current figure

```{r figure-2, eval=FALSE}
```

```{r show-figure-2, echo=FALSE, fig.cap="Figure 2. TE diversity and compositional ordination."}
knitr::include_graphics(
  figure_path("Figure_2_te_diversity_and_ordination.png"),
  rel_path = FALSE
)
```

## Repeat landscape

The production plot selects the six reported categories with the greatest
total aligned bases, combines the remainder as `Other`, zero-fills every
species/category/divergence-bin combination, and gives all 34 species equal
weight in the across-species curves. The young-repeat summary uses bins below
5% divergence.

```{r repeat-support}
repeat_landscape <- read_supp("S12")
repeat_inventory <- read_supp("S13")

stopifnot(
  n_distinct(repeat_landscape$species) == 34L,
  nrow(repeat_inventory) == 34L
)

repeat_inventory %>%
  arrange(desc(aligned_hit_bp)) %>%
  show_table(rows = 12, columns = 4)
```

The full 13,371-row landscape is available as
[S12 CSV](datasets/Supplementary_Data_S12_repeatmasker_divergence_landscape.csv).

```{r figure-3, eval=FALSE}
```

```{r show-figure-3, echo=FALSE, fig.cap="Figure 3. RepeatMasker divergence landscape."}
knitr::include_graphics(
  figure_path("Figure_3_repeat_divergence_landscape.png"),
  rel_path = FALSE
)
```

## LTR terminal:internal depth proxy

### Gate and IQR replay

An eligible element is at least 3,000 bp long, has an exact TEsorter
annotation, is LTR/Gypsy, and has at least five whitespace-delimited
`DOMAIN|MODEL` annotations. The essential filter is inclusive and is computed
within species over the 408 eligible elements:

\[
Q1 - 1.5\,IQR \leq T:I \leq Q3 + 1.5\,IQR
\]

The independent coverage gate requires positive depth across at least 80% of
positions separately in the left LTR, right LTR, and internal region. The
primary branch is the intersection of IQR retention and that coverage gate.

```{r ltr-filter-replay}
ltr_elements <- read_supp("S14")
ltr_species <- read_supp("S15")
ltr_coverage <- read_supp("S16")
ltr_excluded <- read_supp("S17")

count_tesorter_domains <- function(value) {
  text <- trimws(as.character(value))
  if (is.na(value) || text == "" || tolower(text) == "none") return(0L)
  length(strsplit(text, "[[:space:]]+")[[1]])
}

ltr_check <- ltr_elements %>%
  mutate(
    recomputed_domain_count = vapply(Domains, count_tesorter_domains, integer(1)),
    recomputed_domain_gate =
      .data[["element length"]] >= 3000 &
      tesorter_annotation_found &
      Order == "LTR" &
      Superfamily == "Gypsy" &
      recomputed_domain_count >= 5
  )

fences <- ltr_check %>%
  filter(recomputed_domain_gate) %>%
  group_by(species) %>%
  summarise(
    q1 = quantile(ratio_terminal_internal_all_positions, 0.25),
    q3 = quantile(ratio_terminal_internal_all_positions, 0.75),
    iqr = q3 - q1,
    lower = q1 - 1.5 * iqr,
    upper = q3 + 1.5 * iqr,
    .groups = "drop"
  )

ltr_check <- ltr_check %>%
  left_join(fences, by = "species") %>%
  mutate(
    recomputed_iqr_keep = recomputed_domain_gate &
      ratio_terminal_internal_all_positions >= lower &
      ratio_terminal_internal_all_positions <= upper,
    recomputed_coverage =
      left_ltr_positive_coverage_fraction >= 0.8 &
      right_ltr_positive_coverage_fraction >= 0.8 &
      internal_positive_coverage_fraction >= 0.8,
    recomputed_primary = recomputed_iqr_keep & recomputed_coverage
  )

stopifnot(
  nrow(ltr_elements) == 1086L,
  all(ltr_check$recomputed_domain_count == as.integer(ltr_check$domain_count)),
  all(ltr_check$recomputed_domain_gate == as.logical(ltr_check$domain_gate_keep)),
  all(ltr_check$recomputed_iqr_keep == as.logical(ltr_check$iqr_filter_keep)),
  all(ltr_check$recomputed_coverage == as.logical(ltr_check$coverage_ge_80pct)),
  all(ltr_check$recomputed_primary == as.logical(ltr_check$primary_ltr_element)),
  sum(ltr_check$recomputed_domain_gate) == 408L,
  sum(ltr_check$recomputed_iqr_keep) == 381L,
  sum(ltr_check$recomputed_domain_gate & ltr_check$recomputed_coverage) == 407L,
  sum(ltr_check$recomputed_primary) == 380L,
  n_distinct(ltr_check$species[ltr_check$recomputed_primary]) == 30L
)

primary_ltr_species <- ltr_species %>%
  filter(
    analysis_branch == "primary_iqr_filtered_coverage_ge_80pct",
    ratio_median > 0
  ) %>%
  select(
    species, n_elements, ratio_median,
    median_bootstrap_ci_low, median_bootstrap_ci_high,
    fraction_ratio_gt_1,
    median_terminal_positive_coverage_fraction,
    median_internal_positive_coverage_fraction
  ) %>%
  arrange(desc(ratio_median))

stopifnot(nrow(primary_ltr_species) == 30L)
primary_ltr_species %>% show_table(rows = 12, columns = 8)
```

The 1,086-row audit table remains intact; filtering changes branch membership
instead of deleting evidence. No canonical between-species omnibus test has
been finalized for this corrected 380-element branch, so none is added here.

```{r figure-4, eval=FALSE}
```

```{r show-figure-4, echo=FALSE, fig.cap="Figure 4. LTR terminal:internal mapping/deletion-footprint proxy."}
knitr::include_graphics(
  figure_path("Figure_4_ltr_terminal_internal_proxy.png"),
  rel_path = FALSE
)
```

## Cell and nucleus morphology

The object table contains 1,152 frozen reviewed cell-nucleus pairs across 24
species; the report does not silently refilter it. Physical area uses the fixed conversion
\(0.0144\;\mu m^2\) per pixel. Species point estimates are medians of the
accepted reviewed pairs.

```{r morphology-replay}
morphology_objects <- read_supp("S18")
morphology_species <- read_supp("S19")
morphology_audit <- read_supp("S20")

recomputed_morphology <- morphology_objects %>%
  group_by(species) %>%
  summarise(
    n_size_cells = n(),
    n_size_images = n_distinct(filename),
    n_size_specimens = n_distinct(specimen_id),
    cell_area_um2 = median(cell_area_um2),
    nucleus_area_um2 = median(nuc_area_um2),
    .groups = "drop"
  ) %>%
  arrange(species)

observed_morphology <- morphology_species %>%
  select(
    species, n_size_cells, n_size_images, n_size_specimens,
    cell_area_um2, nucleus_area_um2
  ) %>%
  arrange(species)

stopifnot(
  nrow(morphology_objects) == 1152L,
  nrow(morphology_species) == 24L,
  all(abs(morphology_objects$cell_area_um2 - morphology_objects$cell_area_px * 0.0144) < 1e-9),
  all(abs(morphology_objects$nuc_area_um2 - morphology_objects$nuc_area_px * 0.0144) < 1e-9),
  isTRUE(all.equal(recomputed_morphology, observed_morphology, tolerance = 1e-9))
)

morphology_species %>%
  arrange(desc(cell_area_um2)) %>%
  show_table(rows = 12, columns = 12)
```

Seven species are represented by one animal. Intervals are conditional
paired-object intervals for the observed reviewed data; they are not
population-level among-animal uncertainty intervals.

```{r figure-5, eval=FALSE}
```

```{r show-figure-5, echo=FALSE, fig.cap="Figure 5. Reviewed upper-tail cell and nucleus morphology."}
knitr::include_graphics(
  figure_path("Figure_5_reviewed_cell_nucleus_morphology.png"),
  rel_path = FALSE
)
```

## Relative nuclear IOD

Optical density is defined per nuclear-mask pixel from background and observed
intensity; nuclear IOD is the sum of optical density over the mask, not mean
optical density. The primary species estimator is the arithmetic mean of the
within-image median IOD values. Dividing the 24 species estimates and their
interval bounds by the Path24 median anchor produces the relative index:

\[
IOD_{relative,s} = \frac{\operatorname{mean}_{image}
  (\operatorname{median}_{nucleus}(IOD))_s}
  {\operatorname{median}_{species}(IOD_{equal-image})}
\]

A separate descriptive scale uses the published rounded pooled *D. fuscus*
assembly estimate. The declared conversion is:

\[
G_{reference,1C} = \frac{16.1\ \mathrm{Gbp}}
  {0.978\ \mathrm{Gbp\ pg}^{-1}} = 16.462167689\ \mathrm{pg}
\]

Process_413/specimen 32469 and Process_414/specimen 32470 were author-confirmed
as the *D. fuscus* standards and were stained in the same experimental runs as
the unknowns. Their exact standard-to-target run map has not been recovered,
and their median IOD values differ by approximately 41%. The picogram fields
are therefore an assembly-derived conditional sensitivity scale, not an
independently validated absolute C-value; `relative_iod_index` remains the
primary phenotype.

```{r iod-replay}
iod_objects <- read_supp("S21")
iod_species <- read_supp("S22")
iod_images <- read_supp("S23")
iod_balance <- read_supp("S24")
iod_diagnostics <- read_supp("S25")

recomputed_iod <- iod_images %>%
  group_by(species) %>%
  summarise(
    iod_equal_image_estimate = mean(image_median_iod),
    n_images = n_distinct(filename),
    n_specimens = n_distinct(specimen_id),
    .groups = "drop"
  )
iod_anchor <- median(recomputed_iod$iod_equal_image_estimate)
recomputed_iod <- recomputed_iod %>%
  mutate(relative_iod_index = iod_equal_image_estimate / iod_anchor) %>%
  arrange(species)

observed_iod <- iod_species %>%
  select(
    species, iod_equal_image_estimate, n_images, n_specimens,
    relative_iod_index
  ) %>%
  arrange(species)

recomputed_balance <- with(
  iod_balance,
  max_abs_standardized_mean_difference <= 0.10 & max_ks_distance <= 0.25
)

stopifnot(
  nrow(iod_objects) == 805L,
  nrow(iod_images) == 51L,
  nrow(iod_species) == 24L,
  abs(iod_anchor - 950.176112815) < 1e-8,
  isTRUE(all.equal(recomputed_iod, observed_iod, tolerance = 1e-8)),
  identical(
    recomputed_balance,
    as.logical(iod_balance$passes_prespecified_balance_thresholds)
  ),
  sum(!recomputed_balance) == 4L
)

iod_species %>%
  select(
    species, n_nuclei, n_images, n_specimens,
    relative_iod_index, relative_iod_ci_low, relative_iod_ci_high,
    estimate_support, sample_size_support, quality_match_status
  ) %>%
  arrange(desc(relative_iod_index)) %>%
  show_table(rows = 12, columns = 10)
```

The quality-balance rule is `max |SMD| <= 0.10` and `max KS <= 0.25`.
The four failures remain visible in both the data and Figure 6; they are not
silently removed.

```{r iod-quality-table}
iod_balance %>%
  arrange(passes_prespecified_balance_thresholds, desc(max_ks_distance)) %>%
  show_table(rows = 12, columns = 10)
```

<div class="callout warning">
Figure 6 includes frozen pairwise Pagel-lambda PGLS sensitivities. It is not an
absolute-genome-size analysis, and its fitted lines should not be read as
causal relationships.
</div>

```{r figure-6, eval=FALSE}
```

```{r show-figure-6, echo=FALSE, fig.cap="Figure 6. Relative IOD, nucleus area, and cell area sensitivities."}
knitr::include_graphics(
  figure_path("Figure_6_relative_iod_pairwise_pgls.png"),
  rel_path = FALSE
)
```

## Integrated species views

`S40` is the exact 21-species TE34-by-Path24 overlap. The join is at species
level; genomic and microscopy measurements are not from the same animal.

```{r integrated-data}
te_iod_overlap <- read_supp("S40")
organismal_traits <- read_csv(
  file.path(
    ROOT, "analyses", "03_phylogenetic_path", "data",
    "organismal_traits.csv"
  ),
  show_col_types = FALSE
)

stopifnot(
  nrow(te_iod_overlap) == 21L,
  n_distinct(te_iod_overlap$species) == 21L,
  nrow(organismal_traits) == 37L,
  n_distinct(organismal_traits$species) == 37L
)

te_iod_overlap %>%
  select(
    species, scientific_name, shannon_entropy, gini_simpson,
    relative_iod_index, relative_iod_ci_low, relative_iod_ci_high,
    manual_review_inclusion
  ) %>%
  show_table(rows = 12, columns = 8)
```

The expert reproductive-strategy categories are retained as vetted organismal
traits, but the interpretation was supplied after review of the initial PCA.
Any overlay is therefore post hoc and descriptive, not a confirmatory test.

```{r organismal-traits-table}
organismal_traits %>%
  select(
    species, max_svl_mm, development_mode, reproductive_strategy,
    aquaticity_index, microhabitat_class,
    reproductive_strategy_source_id, organismal_source_ids
  ) %>%
  show_table(rows = 12, columns = 8)
```

```{r figure-s1, eval=FALSE}
```

```{r show-figure-s1, echo=FALSE, fig.cap="Supplementary Figure S1. Descriptive TE diversity and relative-IOD overlap."}
knitr::include_graphics(
  figure_path("Figure_S1_te_relative_iod_overlap21.png"),
  rel_path = FALSE
)
```

## Existing frozen statistical sensitivities

<div class="callout warning">
This tab exposes **existing frozen exploratory outputs** so collaborators can
inspect them. It does not promote them to confirmatory manuscript claims. The
report does not run the previously discussed ad hoc power, omnibus, or
univariate phylogenetic-signal calculations. No final between-species LTR test
or diversity-index group test is added here.
</div>

The Path24 traits are log10-transformed and sample-standardized for the current
three-trait analysis. Pairwise PGLS estimates Pagel's lambda by maximum
likelihood. The implemented model contains relative IOD, nucleus area, and cell
area only; the possible integrated TE/SVL path model has not been specified or
fitted.

```{r frozen-statistical-tables}
pairwise_pgls <- read_supp("S26")
path_models <- read_supp("S30")
path_stability <- read_supp("S32")
path_simulation <- read_supp("S36")
path_dsep <- read_supp("S37")

pairwise_pgls %>%
  select(
    comparison, spearman_rho, p_value_descriptive_only,
    pearson_r, pearson_p_value_descriptive_only,
    pgls_standardized_beta, pgls_ci_low, pgls_ci_high,
    pgls_p_value, pagel_lambda, n_species
  ) %>%
  show_table(rows = 3, columns = 11)

path_models %>%
  filter(analysis_type == "primary") %>%
  arrange(delta_CICc) %>%
  show_table(rows = 10, columns = 12)
```

The complete frozen outputs S26–S37 are linked in the Downloads tab. Their
code authority is
[`analyses/03_phylogenetic_path/run_path_analysis.R`](../analyses/03_phylogenetic_path/run_path_analysis.R).

The formal TE-PCA audit remains separate from the descriptive S08–S11 tables:

```{r te-formal-audit}
te_audit_dir <- file.path(ROOT, "analyses", "01_transposable_elements", "data")
te_cluster_diagnostics <- read_csv(
  file.path(te_audit_dir, "te_pca_clustering_diagnostics.csv"),
  show_col_types = FALSE
)
te_cluster_stability <- read_csv(
  file.path(te_audit_dir, "te_pca_clustering_stability.csv"),
  show_col_types = FALSE
)
te_trait_tests <- read_csv(
  file.path(te_audit_dir, "te_pca_trait_association_tests.csv"),
  show_col_types = FALSE
)
te_phylogenetic_tests <- read_csv(
  file.path(te_audit_dir, "te_pca_phylogenetic_signal_tests.csv"),
  show_col_types = FALSE
)

te_cluster_diagnostics %>%
  select(any_of(c(
    "candidate_id", "representation", "k", "cluster_sizes",
    "mean_silhouette", "gap_primary_selected_k", "accepted"
  ))) %>%
  show_table(rows = 12, columns = 8)

te_phylogenetic_tests %>% show_table(rows = 6, columns = 10)
```

No discrete PCA cluster is accepted. Tree-based results remain provisional
until the collaborator tree's publication/calibration provenance is supplied.
The exact replay command is `make te-pca`; it is intentionally not rerun while
knitting this report.

```{r figure-7, eval=FALSE}
```

```{r show-figure-7, echo=FALSE, fig.cap="Figure 7. Exploratory Path24 model and simulation sensitivities."}
knitr::include_graphics(
  figure_path("Figure_7_exploratory_phylogenetic_path_sensitivity.png"),
  rel_path = FALSE
)
```

Auxiliary DAG reference (unnumbered): the current Path24 analysis scores 10
testable Markov-equivalence classes and the auxiliary image below shows one
representative DAG for each class. The saturated 6-member class is omitted
because it has no d-separation claim and was not scored. Relative nuclear IOD
is an image-derived relative phenotype, not absolute genome size, and the
arrows shown here are representative only, not uniquely oriented causal claims.

```{r figure-7-dag-reference, echo=FALSE}
knitr::include_graphics(
  figure_path("Path_DAG_equivalence_class_reference.png"),
  rel_path = FALSE
)
```

## Downloads, provenance, and session

### Controlled CSV downloads

Share this HTML together with the `Publication/datasets/` directory if
collaborators need the linked CSVs.

```{r download-links, results='asis'}
for (section in unique(manifest$analysis_section)) {
  cat("#### ", section, "\n\n", sep = "")
  rows <- manifest %>% filter(analysis_section == section)
  for (index in seq_len(nrow(rows))) {
    cat(
      "- [", rows$supplement_id[[index]], ": ", rows$title[[index]],
      "](datasets/", rows$filename[[index]], ") — ",
      rows$release_status[[index]], "; ", rows$rows[[index]], " rows × ",
      rows$columns[[index]], " columns\n",
      sep = ""
    )
  }
  cat("\n")
}
```

The complete field dictionary is
[COLUMN_INVENTORY.csv](datasets/COLUMN_INVENTORY.csv), and the hash-bound index
is [DATASET_MANIFEST.csv](datasets/DATASET_MANIFEST.csv).

### Report and input fingerprints

```{r report-fingerprints}
tibble(
  artifact = c(
    "Publication/Desmognathus_study_data.Rmd",
    "Publication/datasets/DATASET_MANIFEST.csv",
    "Publication/figures/FIGURE_MANIFEST.csv",
    "scripts/publication/build_gbe_figures.R"
  ),
  sha256 = vapply(
    c(
      REPORT_SOURCE,
      file.path(DATA, "DATASET_MANIFEST.csv"),
      file.path(FIGURES, "FIGURE_MANIFEST.csv"),
      file.path(ROOT, "scripts", "publication", "build_gbe_figures.R")
    ),
    sha256_file,
    character(1)
  )
)
```

### Session information

```{r session-information}
sessionInfo()
```
