---
title: "Desmognathus figure review"
subtitle: "Manifest-driven visual critique gallery"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
    theme: readable
    highlight: tango
---

<style>
:root {
  --review-ink: #1a1a1a;
  --review-muted: #5f6368;
  --review-line: #b8b8b8;
  --review-paper: #ffffff;
  --review-ground: #f3f1ed;
  --review-warning: #fff3df;
  --review-warning-line: #c56f13;
  --review-info: #eef6f8;
  --review-info-line: #247b86;
}

body {
  color: var(--review-ink);
  background: var(--review-ground);
  font-family: Arial, Helvetica, sans-serif;
}

.main-container {
  max-width: 1600px !important;
  padding-bottom: 5rem;
}

h1, h2, h3, h4, .title, .subtitle {
  color: var(--review-ink);
  font-family: Arial, Helvetica, sans-serif;
}

.gallery-summary,
.review-checklist,
.review-callout,
.figure-card {
  background: var(--review-paper);
  border: 1px solid var(--review-line);
}

.gallery-summary,
.review-checklist,
.review-callout {
  margin: 1.25rem 0;
  padding: 1rem 1.2rem;
}

.review-callout {
  border-left-width: 6px;
}

.review-callout.warning {
  background: var(--review-warning);
  border-left-color: var(--review-warning-line);
}

.review-callout.info {
  background: var(--review-info);
  border-left-color: var(--review-info-line);
}

.review-checklist ul {
  columns: 2 28rem;
  column-gap: 2.5rem;
  list-style: none;
  margin: 0;
  padding: 0;
}

.review-checklist li {
  break-inside: avoid;
  margin: 0 0 0.75rem;
  padding-left: 1.6rem;
  position: relative;
}

.review-checklist li::before {
  content: "\2610";
  left: 0;
  position: absolute;
  top: -0.05rem;
}

.figure-card {
  margin: 2rem 0 3rem;
  padding: 1.25rem 1.4rem 1.4rem;
}

.figure-card h2 {
  border: 0;
  margin: 0 0 0.75rem;
  padding: 0;
}

.figure-meta {
  display: flex;
  flex-wrap: wrap;
  gap: 0.45rem 0.65rem;
  margin-bottom: 1rem;
}

.figure-meta span {
  background: #f4f4f4;
  border: 1px solid #d5d5d5;
  border-radius: 2px;
  color: #343434;
  font-size: 0.9rem;
  padding: 0.2rem 0.45rem;
}

.figure-frame {
  border-bottom: 1px solid #d7d7d7;
  border-top: 1px solid #d7d7d7;
  margin: 0 -0.2rem 1rem;
  overflow-x: auto;
  padding: 1.25rem 0.2rem;
  text-align: center;
}

.figure-frame img {
  display: block;
  height: auto;
  margin: 0 auto;
}

.figure-copy {
  display: grid;
  gap: 1rem 1.5rem;
  grid-template-columns: minmax(0, 3fr) minmax(18rem, 2fr);
}

.figure-copy p {
  margin: 0;
}

.figure-copy .alt-text {
  color: var(--review-muted);
}

code {
  color: #252525;
  white-space: pre-wrap;
}

@media (max-width: 850px) {
  .figure-copy {
    grid-template-columns: 1fr;
  }
}
</style>

```{r review-setup, echo=FALSE}
suppressPackageStartupMessages({
  library(readr)
})

ROOT <- normalizePath(getwd(), mustWork = TRUE)
FIGURE_DIR <- file.path(ROOT, "Publication", "figures")
MANIFEST_PATH <- file.path(FIGURE_DIR, "FIGURE_MANIFEST.csv")
AUXILIARY_DAG_BASENAME <- "Path_DAG_equivalence_class_reference.png"
AUXILIARY_DAG_PATH <- file.path(FIGURE_DIR, AUXILIARY_DAG_BASENAME)
AUXILIARY_DAG_RELATIVE_PATH <- file.path("figures", AUXILIARY_DAG_BASENAME)

figure_manifest <- read_csv(
  MANIFEST_PATH,
  show_col_types = FALSE,
  progress = FALSE
)

required_columns <- c(
  "figure_id", "basename", "width_mm", "height_mm", "final_font_family",
  "minimum_target_text_pt", "datasets", "release_status", "legend", "alt_text",
  "pdf", "png", "tiff"
)

if (!all(required_columns %in% names(figure_manifest))) {
  stop(
    "Figure manifest is missing required columns: ",
    paste(setdiff(required_columns, names(figure_manifest)), collapse = ", ")
  )
}
if (nrow(figure_manifest) == 0L || anyDuplicated(figure_manifest$figure_id)) {
  stop("Figure manifest must contain unique, ordered figure records")
}
if (any(!is.finite(figure_manifest$width_mm)) || any(figure_manifest$width_mm <= 0)) {
  stop("Every figure must declare a positive final width")
}

figure_paths <- file.path(FIGURE_DIR, figure_manifest$png)
if (any(!file.exists(figure_paths))) {
  stop(
    "Missing review PNG files: ",
    paste(basename(figure_paths[!file.exists(figure_paths)]), collapse = ", ")
  )
}

maximum_width_mm <- max(figure_manifest$width_mm)
figure_manifest$review_width_percent <-
  100 * figure_manifest$width_mm / maximum_width_mm

escape_html <- function(value, attribute = FALSE) {
  htmltools::htmlEscape(as.character(value), attribute = attribute)
}

format_inline <- function(value) {
  escaped <- escape_html(value)
  gsub("`([^`]+)`", "<code>\\1</code>", escaped, perl = TRUE)
}

figure_one <- figure_manifest[figure_manifest$figure_id == "Figure_1", , drop = FALSE]
figure_one_evidence <- if (nrow(figure_one) == 1L) {
  paste(
    figure_one$datasets,
    figure_one$legend,
    figure_one$alt_text,
    figure_one$basename,
    collapse = " "
  )
} else {
  ""
}
figure_1_has_phylogeny <- nrow(figure_one) == 1L && grepl(
  "S38|S39|phylogen|dated[ -]tree|time-calibrated",
  figure_one_evidence,
  ignore.case = TRUE,
  perl = TRUE
)
```

<div class="gallery-summary">
This page is a visual-review surface, not a second figure authority. It reads
`Publication/figures/FIGURE_MANIFEST.csv` in row order and displays the exact
current review PNG for every record. Declared final widths are preserved
proportionally, so a 120-mm figure appears narrower than a 185-mm figure.
</div>

```{r review-status, echo=FALSE, results='asis'}
cat(
  sprintf(
    "<p><strong>Current gallery:</strong> %d manifest figures; widest declared figure: %.0f mm.</p>\n",
    nrow(figure_manifest),
    maximum_width_mm
  )
)

if (figure_1_has_phylogeny) {
  cat(paste0(
    '<div class="review-callout info"><strong>Figure 1 includes the Path24 dated phylogeny.</strong> ',
    'The tree is shown with the study-design and genomic-QC material so its time scale, ',
    'taxon labels, spacing, and relationship to the other panels can be reviewed together.</div>\n'
  ))
} else {
  cat(paste0(
    '<div class="review-callout info"><strong>Figure 1 phylogeny status:</strong> ',
    'the current manifest does not yet declare a Path24 dated-phylogeny panel. ',
    'This notice will change automatically when Figure 1 records that material.</div>\n'
  ))
}
```

<div class="review-callout warning">
<strong>Unresolved tree provenance.</strong> The focal dated tree was supplied
directly by Alex Pyron, but its publication or archive identifier, calibration method,
and branch-length provenance remain unresolved. Review the displayed
time scale as supplied; do not interpret this gallery as independent validation
of the calibration.
</div>

# Critique checklist

<div class="review-checklist">
<ul>
<li><strong>Narrative and material:</strong> Does every panel earn its place and support the intended manuscript claim?</li>
<li><strong>Cross-figure coherence:</strong> Do repeated concepts use the same visual grammar without implying false equivalence?</li>
<li><strong>Typography:</strong> Are font family, size, weight, italics, symbols, and panel tags consistent and legible at final size?</li>
<li><strong>Color:</strong> Are hues meaningful, high-contrast, color-vision accessible, and reinforced by shape or line type?</li>
<li><strong>Marks and uncertainty:</strong> Are points, lines, intervals, raw observations, and summary estimates visually distinct?</li>
<li><strong>Axes and labels:</strong> Are units, transformations, abbreviations, tick density, and scientific names clear?</li>
<li><strong>Spacing:</strong> Are gutters, panel proportions, margins, legend placement, and whitespace balanced?</li>
<li><strong>Reference-paper fit:</strong> Do framing, density, hierarchy, and annotation feel like one visual family?</li>
<li><strong>Legend and alt text:</strong> Can each figure be understood without manuscript body text?</li>
<li><strong>Phylogeny:</strong> Are topology, tip order, time direction, scale, and provenance boundary immediately clear?</li>
</ul>
</div>

# Figure gallery

```{r manifest-gallery, echo=FALSE, results='asis'}
for (index in seq_len(nrow(figure_manifest))) {
  row <- figure_manifest[index, , drop = FALSE]
  figure_label <- gsub("_", " ", row$figure_id[[1]], fixed = TRUE)
  relative_png <- file.path("figures", row$png[[1]])
  width_percent <- row$review_width_percent[[1]]

  cat(
    sprintf(
      '<section class="figure-card" data-figure-id="%s">\n',
      escape_html(row$figure_id[[1]], attribute = TRUE)
    )
  )
  cat(sprintf("<h2>%s</h2>\n", escape_html(figure_label)))
  cat('<div class="figure-meta">\n')
  cat(sprintf(
    "<span>Final size: %.0f × %.0f mm</span>\n",
    row$width_mm[[1]],
    row$height_mm[[1]]
  ))
  cat(sprintf(
    "<span>Font: %s</span>\n",
    escape_html(row$final_font_family[[1]])
  ))
  cat(sprintf(
    "<span>Target text: ≥%g pt</span>\n",
    row$minimum_target_text_pt[[1]]
  ))
  cat(sprintf(
    "<span>Datasets: %s</span>\n",
    escape_html(row$datasets[[1]])
  ))
  cat(sprintf(
    "<span>Status: <code>%s</code></span>\n",
    escape_html(row$release_status[[1]])
  ))
  cat("</div>\n")

  cat('<div class="figure-frame">\n')
  cat(sprintf(
    paste0(
      '<img src="%s" alt="%s" ',
      'style="width: %.3f%%; max-width: 100%%;" />\n'
    ),
    escape_html(relative_png, attribute = TRUE),
    escape_html(row$alt_text[[1]], attribute = TRUE),
    width_percent
  ))
  cat("</div>\n")

  cat('<div class="figure-copy">\n')
  cat(sprintf(
    '<p class="legend"><strong>Legend.</strong> %s</p>\n',
    format_inline(row$legend[[1]])
  ))
  cat(sprintf(
    '<p class="alt-text"><strong>Alt text.</strong> %s</p>\n',
    format_inline(row$alt_text[[1]])
  ))
  cat("</div>\n</section>\n")
}
```

<div class="review-callout info">
Record critiques by figure ID and panel letter. Changes belong in the shared
ggplot theme or production figure builder, not in this gallery or its embedded
images.
</div>

# Auxiliary path-DAG reference

```{r auxiliary-dag-reference, echo=FALSE, results='asis'}
cat(
  paste0(
    '<section class="figure-card" data-figure-id="Auxiliary_Path_DAG_reference">',
    '<h2>Auxiliary Path DAG reference</h2>',
    '<div class="figure-meta">',
    '<span>Status: <code>auxiliary_review_only</code></span>',
    '<span>Asset: <code>', escape_html(AUXILIARY_DAG_BASENAME), "</code></span>",
    "</div>",
    '<div class="review-callout info"><strong>Not a numbered manuscript figure.</strong> ',
    'This auxiliary review-only graphic shows one representative for each of the 10 testable equivalence classes. ',
    'The saturated 6-member class is not shown and remains unscored because it has no d-separation claim. ',
    'The current Path24 relative IOD-only analysis remains exploratory and does not orient a unique causal arrow.</div>'
  )
)

if (file.exists(AUXILIARY_DAG_PATH)) {
  cat('<div class="figure-frame">\n')
  cat(sprintf(
    paste0(
      '<img src="%s" alt="%s" ',
      'style="width: 100%%; max-width: 100%%;" />\n'
    ),
    escape_html(AUXILIARY_DAG_RELATIVE_PATH, attribute = TRUE),
    escape_html(
      paste(
        "Auxiliary review-only path-DAG reference showing one representative",
        "for each of the 10 testable equivalence classes, with the saturated",
        "6-member class identified as unscored for the current Path24 relative IOD-only analysis."
      ),
      attribute = TRUE
    )
  ))
  cat("</div>\n")
} else {
  cat(
    paste0(
      '<div class="review-callout warning"><strong>Auxiliary DAG image pending.</strong> ',
      'Expected <code>', escape_html(AUXILIARY_DAG_BASENAME),
      "</code> under <code>Publication/figures/</code> before the self-contained review HTML is refreshed.</div>\n"
    )
  )
}

cat(
  paste0(
    '<div class="figure-copy">',
    '<p class="legend"><strong>Boundary.</strong> One representative for each of the 10 testable equivalence classes is shown for review. ',
    'The saturated 6-member class is not shown and remains unscored because it has no d-separation claim.</p>',
    '<p class="alt-text"><strong>Scope.</strong> Current Path24 relative IOD only; no TE, SVL, or unique causal-arrow claim is introduced by this auxiliary panel.</p>',
    "</div></section>\n"
  )
)
```
